A Gentle Introduction into Structural Causal Models

### 2021-05-07 topic_1_firstdraft.Rmd
topic 1 table

@@

@@

Independence

>>>>

Aside from path diagrams also introduced graph rules for writing down relationships.

@@

Causal assumptions are encoded in missing links.

@@

Causal assumptions differentiate causal models from association learning methods.

 | Association-based Concepts |  Causal Concepts
|--------------------------|------------------------|
| Correlation              | Randomization  
| Regression               | Confounding
| Conditional Independence | Disturbance
| Likelihood               | Error Terms      
| Odds Ratio               | Structural Coefficients       
| Propensity Score         | Spurious Correlation

Table: Concepts in Causality and Association concepts

These causal concepts are not expressible based on distribution functions/statistical associations. (Pearl 2010)

Disturbance in SCM: Correlated and causal factor ; responsible for variation
Disturbance in Regression: Uncorrelated

Causal assumption not testable (e.g. $Cov(U_a, U_b)=0$).
d-seperation to test assumptions in totality (cannot make assumptions in isolation).

exploit invariant characteristics of SEM whithout committing to shape. 
structural if function autonomous and invariant to change in form of other functions.
